---
title: "Seaworld"
author: "Yifeng Huang"
date: "August 16, 2018"
output: pdf_document
---
```{r lib,echo=FALSE,warning=F}
library(readxl)
library(MASS)
library(car)
library(corrplot)
library(psych)
library(ggplot2)
library(tidyr)
```

#Data Clean


##Overview
```{r,warning=F}
#data import
seawatch.ori<-read_excel("~/MSBA notes/Business Stats/Seawatch C w blanks-1.xlsx")

# we delete CNVHRS, Notes, City and Zip code. 
seawatch<-seawatch.ori[,3:20]
seawatch<-seawatch[,-2] 
# overview
head(seawatch)
describe(seawatch)

ggplot(gather(seawatch), aes(value)) + 
    geom_histogram(bins = 10) + 
    facet_wrap(~key, scales = 'free_x')
## Note that VISIT and LST are categorical variables

## Convert numberic variables to Categorical
seawatch$VISIT<-as.factor(seawatch$VISIT)
seawatch$LST<-as.factor(seawatch$LST)
```

##Missing Values
```{r}
# number of NA's
nrow(seawatch)-nrow(na.omit(seawatch))

## Since there are only 34 rows containing na's we can simply delete it
seawatch<-na.omit(seawatch)
```


##Other strange observation


###CART, REAG, and ANDR


- For some observations, the number of votes is bigger than total population. Due to the high correlation (0.9533607) between POP80 and the sum of those there vote numbers, we can build a model to predict the right population.
```{r}
# scatter plot
plot(GROSS~I(CART/POP80),data = seawatch)

# observations that CART or REAG or ANDR is larger than total population
ex.obs<-seawatch$CART>seawatch$POP80 | seawatch$REAG>seawatch$POP80 | seawatch$REAG>seawatch$POP80

seawatch[ex.obs,c("POP80",'REAG','ANDR')]

# correlation
cor(seawatch$POP80,I(seawatch$CART+seawatch$REAG+seawatch$ANDR))

# plot pop80 and sum of those 3 vote numbers
plot(POP80~I(CART+REAG+ANDR),data = seawatch)

# original sd
sd(seawatch[-ex.obs,]$POP80)

# predictive model
## full model
pop.lm<-lm(POP80~.+I(CART+REAG+ANDR),data = seawatch[-ex.obs,])
## Predictors selection
step(pop.lm,direction = "backward",trace = 0)
## update model
pop.lm<-lm(formula = POP80 ~ GROSS + MOY + CPI + HHMEDI + POVPR + MFGPR + 
    MAGE + CART + REAG + ANDR,data = seawatch[-ex.obs,])

## check multicolinearity
vif(pop.lm)

## drop ANDR
pop.lm<-update(pop.lm,.~.-ANDR)
vif(pop.lm)

## summary and plot of the model
summary(pop.lm)

## drop GROSS, MOY, CPI and HHMEDI
pop.lm<-update(pop.lm,.~.-GROSS-MOY-CPI-HHMEDI)
summary(pop.lm)
par(mfrow=c(2,2))
plot(pop.lm)

## predict the total population
pred<-predict(pop.lm,seawatch[ex.obs,])
seawatch[ex.obs,"POP80"]<-pred
```


###POVPR,COLLPR and MFGPR


- Due to the increasing variance, instead of percentage, we transform those varibles to exact number by mutiplying total population. As a result, the linear correlations is more obvious.
```{r,warning=F}
#Compare variance before and after transformation
par(mfrow=c(2,3))
plot(GROSS~POVPR,data = seawatch)
plot(GROSS~MFGPR,data = seawatch)
plot(GROSS~COLLPR,data = seawatch)
plot(GROSS~I(POVPR*POP80),data = seawatch)
plot(GROSS~I(MFGPR*POP80),data = seawatch)
plot(GROSS~I(COLLPR*POP80),data = seawatch)

# Add POVPP,MFGPP,COLLPP to data
seawatch$POVPP<-seawatch$POVPR*seawatch$POP80
seawatch$MFGPP<-seawatch$MFGPR*seawatch$POP80
seawatch$COLLPP<-seawatch$COLLPR*seawatch$POP80

```


#Modeling


##Training and Testing subsets split
```{r}
set.seed(1024)
train.num<-sample(1:dim(seawatch)[1],round(nrow(seawatch)*0.75))
seawatch.train<-seawatch[train.num,]
seawatch.test<-seawatch[-train.num,]
```



##Predictors Selections
```{r}

# full model
full.lm<-lm(data = seawatch.train,GROSS~.)

# predictors selection
step(full.lm,direction = "backward",trace = 0)

# update model
fit.lm<-lm(formula = GROSS ~ MOY + YR + VISIT + LST + HHMEDI + CART + 
    REAG + ANDR + POVPP + COLLPP, data = seawatch.train)

# check multicolinearity
vif(fit.lm)

# drop ANDR
fit.lm<-update(fit.lm,.~.-ANDR)
vif(fit.lm)

# drop CART
fit.lm<-update(fit.lm,.~.-CART)
vif(fit.lm)


# summary 
summary(fit.lm)

# drop YR,POVPR,PERCAPI
fit.lm<-update(fit.lm,.~.-YR-MOY-HHMEDI-REAG)
summary(fit.lm)

# residual analysis
par(mfrow=c(2,2))
plot(fit.lm)
```


- There exists clear non-constant variance. Also predictors are not normally distributed. As a result, we use the power transformation to modify the model.


##Power Transformation model
```{r}
powerTransform(cbind(seawatch.train$GROSS,seawatch$POVPR,seawatch$COLLPP)~1)

# Thus, we build another model by taking natural log on both sides
new.fit.lm<-lm(formula = log(GROSS) ~ VISIT + LST + log(POVPP) + log(COLLPP), data = seawatch.train)
summary(new.fit.lm)
par(mfrow=c(2,2))
plot(new.fit.lm)
```


- Note that the variance of error is more constant and the predictors are distributed better than the original linear model


##Model based on correlation
```{r}
model2<- lm(GROSS ~ VISIT + POP80 + PERCAPI + MFGPR + REAG + POVPP + COLLPP,data = seawatch.train)
summary(model2)

par(mfrow=c(2,2))
plot(model2)

#multicolinearity check
vif(model2)

```

##Cross Validation
```{r}
#MSE function
MSE<-function(pred,actual){
  return(mean((pred-actual)^2))
}

#predictions based on each model
pred.fit<-predict(fit.lm,newdata = seawatch.test)
pred.fit.new<-predict(new.fit.lm,newdata = seawatch.test)
pred.model2<-predict(model2,newdata = seawatch.test)

#MSE table
data.frame(
  Model=c("fit.lm","fit.powertrans","model2"),
  MSE=c(MSE(pred.fit,seawatch.test$GROSS),MSE(exp(pred.fit.new),seawatch.test$GROSS),MSE(pred.model2,seawatch.test$GROSS))
)
```


- The model with lowest MSE so far is the log-transformation model.




